# Packages

```{r}
#remotes::install_github("hrbrmstr/albersusa")
#install.packages("NbClust")
#install.packages("factoextra")
```

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(sf)
library(NbClust)
library(factoextra)
library(magrittr)
```

### Reading ESRI Shapefiles (Polygonal Vectorial data)

```{r}
counties_sf <- read_sf('../Data/spatial_data/counties_sf.shp')

usa_states <- read_sf('../Data/spatial_data/usa_sf.shp')
```

# ----------------------------------------------

### Vectorial data with country, states and counties boundaries

```{r}
counties_usa <- counties_sf
counties_usa %>% 
  data.frame()
```

```{r}
ggplot() +
  geom_sf(data = usa_states)
```

# ----------------------------------------------

## Example 1: Mesonet stations visualization

```{r}
kansas_counties <-  counties_usa %>% 
  filter(state == 'Kansas')
```

```{r}
ggplot() + 
  geom_sf(data = kansas_counties)
```

```{r}
ggplot() +
    geom_sf(data = usa_states) +
    geom_sf(data = kansas_counties, aes(fill=as.factor(name)),
            show.legend = FALSE) +
    scale_fill_viridis_d() +
    coord_sf(crs = st_crs(4326), 
             xlim = c(-104, -93), ylim = c(36, 41.5))
```

### Load the mesonet weather stations locations files

```{r}
mesonet_sites <- read.csv('../data/spatial_data/KSMeso-Station-Metadata.csv',skip = 2)
mesonet_sites
```

```{r}
mesonet_sites <- mesonet_sites %>% 
  st_as_sf(coords = c("Longitude","Latitude")) %>%
  st_set_crs(st_crs("epsg:4326"))

plot(mesonet_sites)
```

### Use ggplot2 to plot and visualize the information. Add the name of axis X and Y, Title and subtitle

```{r}
ggplot() +
    geom_sf(data = usa_states,color='#343a40',fill='gray',size=0.4) +
    geom_sf(data = kansas_counties, aes(fill=as.factor(name)),
            show.legend = FALSE,alpha=0.5,size=0.2,color='black') + 
    geom_sf(data = mesonet_sites,size=1, color='darkgreen',
            fill='black')+ 
    scale_fill_viridis_d() +
  xlab('Longitude') +
  ylab('Latitude') +
    coord_sf(crs = st_crs(4326), 
             xlim = c(-104, -93), ylim = c(36, 41)) 
```

# ----------------------------------------------

## Example 2: Relationship among pivot irrigation fields, streams and groundwater resources.

```{r}
streams <- read_sf('../Data/spatial_data/kansas_streams_order3_9_webmerc.shp') %>%
  select(GNIS_NAME)

```

```{r}
high_plain_groundwater <- read_sf('../Data/spatial_data/hp_bound2010_wgs84.shp') %>%
  st_set_crs(st_crs("epsg:4326"))

```

```{r}
ggplot() +
  geom_sf(data = usa_states,) +
  geom_sf(data = high_plain_groundwater, fill=NA, color='blue',show.legend = TRUE)  +
    coord_sf(crs = st_crs(4326), 
             xlim = c(-110, -90), ylim = c(30, 45)) 
```

```{r}
ggplot() + 
  geom_sf(data = kansas_counties) +
  geom_sf(data = streams)
```

```{r}
field_irrigated <- read_csv('../Data/spatial_data/field_irrigated_ks.csv') %>%
  select(LONGITUDE,LATITUDE,WELL_USE) %>%
  filter(WELL_USE == 'Irrigation') %>%
  na.exclude() %>%
  st_as_sf(coords=c('LONGITUDE','LATITUDE'))  %>%
  st_set_crs(st_crs("epsg:4326"))
```

```{r}

boundary_kansas <- usa_states %>% 
  filter(name=='Kansas')

ggplot() + 
  geom_sf(data = boundary_kansas) +
  geom_sf(data = streams,color='blue') + 
  geom_sf(data = field_irrigated, color='black', size=0.1)
```

### Clipping high_plain_groundwater layer with Kansas polygon (boundary_kansas)

```{r}

high_plain_groundwater <- high_plain_groundwater  %>%
  st_make_valid()

boundary_kansas <- boundary_kansas %>% 
  st_make_valid()

high_plain_groundwater_ks <- high_plain_groundwater %>% 
  st_intersection(boundary_kansas)

ggplot() + 
  geom_sf(data = boundary_kansas) + 
  geom_sf(data = high_plain_groundwater_ks)

```

### Visualizing all layers at the same time.

```{r}
ggplot() + 
  geom_sf(data = boundary_kansas) +
  #geom_sf(data = streams,color='blue') + 
  geom_sf(data = high_plain_groundwater_ks,fill=NA,color='orange',size=3,alpha=0.7) + 
  geom_sf(data = field_irrigated, color='black', size=0.1,alpha=0.4)


```

# ----------------------------------------------

## Example 3: Grid creation

```{r create grid}

boundary_kansas_proj <- boundary_kansas %>%
  st_set_crs(st_crs("epsg:4326")) %>%
  st_transform(crs=3744) 

grid <- st_make_grid(boundary_kansas_proj,
                     cellsize = 25000,
                     what = "polygons") 

ggplot() + 
  geom_sf(data = boundary_kansas) + 
  geom_sf(data = grid)
```

### Grid clipping

```{r}
grid <- boundary_kansas_proj %>% 
  st_intersection(grid) %>%
  st_set_crs(st_crs("epsg:3744")) %>%
  st_transform(crs=4326) 
```

### Grid visualization

```{r}
ggplot() + 
  geom_sf(data = boundary_kansas) + 
  geom_sf(data = grid)  
```

### Reading terraclimate rasters

```{r}
precipitation <- raster::raster('../Data/spatial_data/pet.tif')
eto <- raster::raster('../Data/spatial_data/aet.tif')
temp_min <- raster::raster('../Data/spatial_data/tmmn.tif')
temp_max <- raster::raster('../Data/spatial_data/tmmx.tif')


df_climate_vars <- raster::stack(precipitation,eto,temp_max,temp_min) %>%
  raster::extract(grid, fun='median') %>%
  cbind(grid)
  

```

### Raster extraction from raster and polygon

```{r}
future::plan(future::multicore(workers = availableCores(constraints = "multicore")))
res <- NbClust(df_climate_vars %>%
                 st_drop_geometry() %>%
                 select(pet,aet, tmmx, tmmn) %>% 
                 tibble::remove_rownames(), 
               diss = NULL, distance = 'euclidean', min.nc = 6, max.nc = 85, 
               method = "kmeans", index = 'all')


```

### Environmental zone delineation with Clustering analysis using Kmeans

```{r}
set.seed(123)
km.res <- kmeans(df_climate_vars %>%
                 st_drop_geometry() %>%
                 select(pet,aet, tmmx, tmmn) %>% 
                 tibble::remove_rownames(), centers = 6, nstart = 25)

df_climate_vars %<>% 
  mutate(zone = as.factor(km.res$cluster))


```

### Environmental Zone visualization

```{r}
ggplot() + 
  geom_sf(aes(fill=zone), data = df_climate_vars)
```

## Resources

-   <https://www.sciencebase.gov/catalog/item/6314061bd34e36012efa397b>

-   
